Data Exploration: IUCN and AquaMaps

### 2016-08-12 coralmaps/coral_maps.Rmd
tracking down coral bathymetry questions

  ### LOICZID cell ID numbers - used in plotting rasters
  loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
  loiczid_raster       <- raster(loiczid_raster_file)
  names(loiczid_raster) <- 'loiczid'

@@

  write_csv(bath_df, file.path(dir_git, 'coralmaps', 'bathy_cells.csv'))
  ### probably want to filter to just 100% 200m+
}

@@

### load iucn spp_cells
coral_files <- list.files(file.path(dir_anx, 'v2016', 'iucn_intersections'), 
                          pattern = 'CORAL*.*',
                          full.names = TRUE)
coral_cells_iucn <- lapply(coral_files, read_csv) %>%
  bind_rows() %>%
  select(iucn_sid, loiczid = LOICZID) %>%
  filter(iucn_sid %in% spp_depth$iucn_sid[spp_depth$paired_map]) %>%
  left_join(bath_df, by = 'loiczid') %>%
  mutate(deep = ifelse(is.na(deep), 0, deep))

@@

### get species list
data_file <- file.path(dir_data, 'spp_list_w_area_trimmed.csv')
spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
  filter(!is.na(popn_category)) %>%   ### filter out unassessed species
  filter(!is.na(sm_perc))             ### filter out NA maps

@@

  ### get species list
  coral_cells_iucn_clipped <- read_csv(file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv')) %>%
    filter(deep == 0) %>%
    mutate(prop_area = 1)

  spp_coralmaps <- read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv')) %>%
    filter(iucn_sid %in% coral_cells_iucn_clipped$iucn_sid) %>%
    select(am_sid, sciname, iucn_sid)

@@

  loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
  loiczid_raster       <- raster(loiczid_raster_file)
  names(loiczid_raster) <- 'loiczid'

@@

    message(sprintf('%s. Processing species iucn_sid: %s \n   am_sid(s):  %s \n   sciname(s): %s', 
                    i, sid, paste(sids$am_sid, collapse = ', '), paste(sids$sciname, collapse = ' ')))
    map_iucn <- get_spp_map(sid, coral_cells_iucn_clipped) %>%
      mutate(iucn_pres = ifelse(area_iucn > 0, 1, 0))
    map_am   <- get_spp_map(sids$am_sid, coral_cells_am) %>%
      mutate(am_pres = ifelse(am_prob >= am_thresh, 1, 0))

@@

  ### combine old and new area dataframes
  spp_coralmaps <- spp_coralmaps %>%
    select(iucn_sid, am_sid, sciname, area_am, 
           area_iucn_clipped = area_iucn, area_overlap_clipped = area_overlap, 
           sm_perc_clipped = sm_perc, sm_range_clipped = sm_range, area_ratio_clipped = area_ratio) %>%
    left_join(read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv')) %>%
                select(am_sid, iucn_sid, 
                       area_iucn_raw = area_iucn, area_overlap_raw = area_overlap, 
                       sm_perc_raw = sm_perc, sm_range_raw = sm_range, area_ratio_raw = area_ratio),
              by = c('iucn_sid', 'am_sid'))

@@

``` {r}

spp_coralmaps <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv')) %>% 
                      mutate(loss_ratio = 100 * area_iucn_clipped/area_iucn_raw)

spp_small_effect <- spp_coralmaps %>%
  filter(loss_ratio > 80) 
### only five with such small effects; one of those did not change at all.

summary(spp_coralmaps$loss_ratio)
  #  Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  # 23.80   52.89   55.48   56.33   60.59  100.00

hist(spp_coralmaps$loss_ratio)
### check raw iucn_area vs clipped - looks proportional; scaling factor due to difference in area calcs?

scatterplot1 <- ggplot(spp_coralmaps, aes(x = area_iucn_raw, y = area_iucn_clipped)) +
  geom_point() +
  geom_abline(color = 'red')
scatterplot1

violplot1 <- ggplot(spp_coralmaps, aes(x = 'sm_range_raw', y = loss_ratio)) +
  ggtheme_plot +
  geom_violin(fill = 'grey80', alpha = .5,
              draw_quantiles = c(.25, .5, .75)) +
  geom_point(position = position_jitter(w = .3), 
             color = '#4d4dac', alpha = .5) +
  geom_violin(fill = NA, draw_quantiles = c(.25, .5, .75)) +
  geom_hline(color = 'grey30', yintercept = 100) +
  scale_y_continuous(limits = c(0, 100)) +
  scale_x_discrete(labels = element_blank()) +
  theme(axis.title.x = element_blank()) +
  labs(y = 'IUCN coral range: (<= 200 m) / (total)')
print(scatterplot1)

### recalc areas and overlaps
iucn_coral_fixed <- spp_coralmaps %>%
  gather(dist, perc, c(sm_perc_raw, sm_perc_clipped)) %>%
  gather(area, ratio, c(area_ratio_raw, area_ratio_clipped)) %>%
  filter(!(str_detect(dist, 'raw') & !(str_detect(area, 'raw')))) %>%
  filter(!(str_detect(dist, 'clip') & !(str_detect(area, 'clip')))) %>%
  mutate(method = ifelse(str_detect(dist, 'clip'), '200 m clip', 'original'))

>>>>

  ### LOICZID cell ID numbers - used in plotting rasters
  loiczid_raster_file  <- file.path(dir_git, 'shiny_am_iucn/data/loiczid_raster.tif')
  loiczid_raster       <- raster(loiczid_raster_file)
  names(loiczid_raster) <- 'loiczid'

@@

  writeRaster(bath_raster, file.path(dir_git, 'coralmaps', 'bathy_cells.tif'))

  write_csv(bath_df, file.path(dir_git, 'coralmaps', 'bathy_cells.csv'))
  # bath_df <- read_csv(file.path(dir_git, 'coralmaps', 'bathy_cells.csv'), col_types = 'dd')
  ### probably want to filter to just 100% 200m+
}

@@

spp_depth <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_depth.csv'))

### load iucn spp_cells
coral_files <- list.files(file.path(dir_anx, 'v2016', 'iucn_intersections'), 
                          pattern = 'CORAL*.*',
                          full.names = TRUE)
coral_cells_iucn <- lapply(coral_files, read_csv) %>%
  bind_rows() %>%
  select(iucn_sid, loiczid = LOICZID) %>%
  # filter(iucn_sid %in% spp_depth$iucn_sid[spp_depth$paired_map]) %>%
  left_join(bath_df, by = 'loiczid') %>%
  mutate(deep = ifelse(is.na(deep), 0, deep))

@@

### get species list
data_file <- file.path(dir_data, 'spp_list_w_area_trimmed.csv')
spp_map_pairs <- read_csv(data_file) %>%
  filter(!is.na(popn_category)) %>%   ### filter out unassessed species
  filter(!is.na(sm_perc))             ### filter out NA maps

@@

  ### get species list
  coral_cells_iucn_shallow <- read_csv(file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv')) %>%
    filter(deep == 0) %>%
    mutate(prop_area = 1)

  spp_coralmaps <- read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv')) %>%
    filter(iucn_sid %in% coral_cells_iucn_shallow$iucn_sid) %>%
    select(am_sid, sciname, iucn_sid)

@@

  loiczid_raster_file  <- file.path(dir_git, 'shiny_am_iucn/data/loiczid_raster.tif')
  loiczid_raster       <- raster(loiczid_raster_file)
  names(loiczid_raster) <- 'loiczid'

@@

    message(sprintf('%s. Processing species iucn_sid: %s \n   am_sid(s):  %s \n   sciname(s): %s', 
                    i, sid, paste(sids$am_sid, collapse = ', '), paste(sids$sciname, collapse = ' ')))
    map_iucn <- get_spp_map(sid, coral_cells_iucn_shallow) %>%
      mutate(iucn_pres = ifelse(area_iucn > 0, 1, 0))
    map_am   <- get_spp_map(sids$am_sid, coral_cells_am) %>%
      mutate(am_pres = ifelse(am_prob >= am_thresh, 1, 0))

@@

  ### combine old and new area dataframes
  spp_coralmaps <- spp_coralmaps %>%
    select(iucn_sid, am_sid, sciname, area_am, 
           area_iucn_shallow = area_iucn, area_overlap_shallow = area_overlap, 
           sm_perc_shallow = sm_perc, sm_range_shallow = sm_range, area_ratio_shallow = area_ratio) %>%
    left_join(read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv')) %>%
                select(am_sid, iucn_sid, 
                       area_iucn_raw = area_iucn, area_overlap_raw = area_overlap, 
                       sm_perc_raw = sm_perc, sm_range_raw = sm_range, area_ratio_raw = area_ratio),
              by = c('iucn_sid', 'am_sid'))

@@

``` {r generate_area_plots}

spp_coral_clip <- read_csv(file.path(dir_git, 'coralmaps', 'iucn_coral_areas.csv')) %>%
  select(iucn_sid, shallow_pct, totarea, deeparea) %>%
  mutate(shallow_pct = 100 * shallow_pct)

spp_small_effect <- spp_coral_clip %>%
  filter(shallow_pct > 80) 
### only five with such small effects; one of those did not change at all.

summary(spp_coral_clip$shallow_pct)
  #  Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  # 23.80   52.89   55.48   56.33   60.59  100.00

scatterplot1 <- ggplot(spp_coral_clip, aes(x = totalarea, y = deeparea)) +
  geom_point(size = 0.25, color = 'black') +
  geom_abline(color = 'red')
scatterplot1

violplot1 <- ggplot(spp_coral_clip, aes(x = 'sm_range_raw', y = shallow_pct)) +
  ggtheme_plot +
  # geom_violin(fill = 'grey80', alpha = .5,
  #             draw_quantiles = c(.25, .5, .75)) +
  # geom_point(position = position_jitter(w = .3), 
  #            size = 0.25,
  #            color = 'black', alpha = 1) +
  geom_violin(fill = NA, size = 0.1, weight = 0.5,
              draw_quantiles = c(seq(.005, .05, .005), seq(.05, .95, .02), seq(.95, 1.0, .005)),
              color = 'black') +
  geom_violin(fill = NA, size = 1,
              draw_quantiles = NULL,
              color = 'white') +
  geom_hline(color = 'grey30', yintercept = 100) +
  scale_y_continuous(limits = c(0, 100)) +
  scale_x_discrete(labels = element_blank()) +
  theme(axis.title.x = element_blank()) +
  labs(y = 'IUCN coral range: (<= 200 m) / (total)')
print(violplot1)

spp_coral_clip2 <- spp_coral_clip %>%
  select(iucn_sid, shallow = shallow_pct) %>%
  arrange(desc(shallow))

spp_coral_clip2$l_to_h <- c(1:nrow(spp_coral_clip))

spp_coral_clip2 <- spp_coral_clip2 %>%
  mutate(deep = 100 - shallow) %>%
  gather(depth, pct_value, c(shallow, deep)) %>%
  mutate(depth = ifelse(depth == 'shallow', '0 - 200 m', '200 m +'))

break_nums <- seq(0, 100, 20)
coral_area_plot <- ggplot(spp_coral_clip2, aes(x = l_to_h, y = pct_value, fill = depth)) +
  ggtheme_plot +
  # geom_linerange(aes(ymin = kept_ratio, ymax = 100),
  #                position = position_dodge(width = 0.5)) +
  geom_area() +
  scale_fill_manual(values = c('0 - 200 m' = 'grey90', '200 m +' = 'grey10')) +
  geom_hline(yintercept = break_nums, color = 'white', size = .25) +
  scale_y_continuous(limits = c(0, 100), breaks = break_nums,
                     labels = paste0(break_nums, '%')) +
  # scale_x_discrete(labels = element_blank()) +
  # theme(axis.title.x = element_blank()) +
  labs(y = 'IUCN % of coral range',
       x = 'Count of coral species')
print(coral_area_plot)

ggsave(plot = coral_area_plot,
       filename = file.path(dir_fig, 'coral_area_plot.png'), device = 'png',
       width = 9, height = 5, units = 'cm', dpi = 600)

``` {r generate_coral_quadmap}

spp_coralmaps <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv'))

### recalc areas and overlaps
iucn_coral_fixed <- spp_coralmaps %>%
  gather(dist, perc, c(sm_perc_raw, sm_perc_shallow)) %>%
  gather(area, ratio, c(area_ratio_raw, area_ratio_shallow)) %>%
  filter(!(str_detect(dist, 'raw') & !(str_detect(area, 'raw')))) %>%
  filter(!(str_detect(dist, 'clip') & !(str_detect(area, 'clip')))) %>%
  mutate(method = ifelse(str_detect(dist, 'clip'), '200 m clip', 'original'))
